Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2019-03-02 documentation/report.tex
Formatting resolutions

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyhf{}
\lfoot{\scriptsize January 2019}
\cfoot{\scriptsize Brendon Matusch---Improving Particle Classification in WIMP Experiments}
\rfoot{\scriptsize Page~\thepage~of~\pageref{LastPage}}

@@

In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) arranged in a hexagonal lattice on the surface of the spherical detector. The randomness severely limits the accuracy of conventional classifiers; the best classifier so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events. Because of physical limitations, simulated data is used for calibration, with a limited amount of real-world experimental data available for testing.

I composed a 28-page research whitepaper \cite{me} about my work on PICO-60, which has been reviewed and approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}. It is currently undergoing peer review for eventual publication in the journal \textit{Computer Physics Communications}.

@@

In the context of PICO-60, my intent was to develop a highly accurate classification algorithm that is tolerant to substantial impurities in calibration data. I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.

@@

Because of the inherently spatial nature of this problem, I hypothesized it would be important to most effectively process the spherical arrangement of PMTs in the DEAP-3600 detector. I evaluated three approaches to this problem (two of which are original), listed below. Techniques \ref{map_entry} and \ref{topological_entry} (Figure \ref{deap_methods}) take advantage of geometric and spatial patterns in PMT activations associated with neck alpha events, by extending the CNN concept for use in spherical and arbitrary topologies.

@@

I evaluated three algorithms for resolution of the unpredictable photon dynamics present in DEAP-3600. Two of the algorithms are original. The three algorithms I developed were evaluated based on:

>>>>

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\fancyhf{}
\lfoot{\scriptsize January 2019}
\cfoot{\scriptsize Improving Particle Classification in WIMP Experiments -- Brendon Matusch}
\rfoot{\scriptsize Page~\thepage~of~\pageref{LastPage}}

@@

In DEAP-3600, photons are absorbed by a wavelength shifting medium and re-emitted in an unpredictable direction, before being detected by one of 255 photomultiplier tubes (PMTs) on the surface of the spherical detector. The randomness severely limits the accuracy of conventional classifiers; the best so far removes 99.6\% of alpha background radiation, while also (undesirably) removing 91.0\% of WIMP events. Because of physical limitations, simulated data is used for calibration, with 30 real-world experimental events available for testing.

I have written a research paper \cite{me} about my work on PICO-60, which has been approved by the PICO collaboration and pre-published at \url{https://arxiv.org/abs/1811.11308}. It is currently undergoing peer review for publication in \textit{Computer Physics Communications}.

@@

In the context of PICO-60, my intent was to develop a highly accurate classification algorithm

that is tolerant to substantial impurities in calibration data. I developed and compared two sets of classification algorithms: supervised learning, and semi-supervised learning.

@@

Because of the inherently spatial nature of this problem, I hypothesized it would be important to most effectively process the arrangement of PMTs in the DEAP-3600 detector: a hexagonal lattice on the surface of a sphere. I evaluated three approaches to this problem (two of which are original), listed below. Techniques \ref{map_entry} and \ref{topological_entry} (Figure \ref{deap_methods}) take advantage of geometric and spatial patterns in PMT activations associated with neck alpha events, by extending the CNN concept for use in spherical and arbitrary topologies.

@@

The three I developed to resolve the problem of unpredictable dynamics in DEAP-3600 were evaluated based on:
